---
title: "DA"
format: html
editor: visual
---
This is my learning summary of the book "Introduction to statistical learning".

## Discriminant analysis

classification problems

### Bayes claasifier

Bayes theoretical boundary that minimizes the classification error $$
P(Y=k|X=x)>P(Y=j|X=x)
$$ (use P（X\|Y=k) and P(Y=k) for all "k"s.) for all j is not equal to k.

### LDA

assumes a common covariance matrix and normal distribution

### QDA

assumes their own nomal distribution's covariance matrix with normal distributions

### Naive Bayes

assumes independence

Useful when p is large(QDA, LDA break down because of too many 2-d plot???)

Gaussian navie bayes assumes each covariance matrix is diagonal

### Why Discriminant Analysis?

(1)Logistic regression is unstable when estimating parameters for well-separated classes(if perfectly, **????infinite coefficients**). LDA does not suffer from it (based on Bayes theo)

(2)When n is small and predictors X is approximately normal in each class, **???LDA is more stable than the logistic regression model.**

(3)**????LDA is popular when we have 2 more response classes because it provides a low-dimensional view of data.**

(4)the dummy variable approach cannot be easily extended to accommodate qualitative responses with more than two levels

#### LDA and Bayes classifier

(1)LDA is less flexible than Bayes classifier(only linear decision boundaries)

(2)Bayes classifier is optimal if true distributions are known(P(X,Y))

(3)LDA assumes normally distributed classes with equal covariance matrices.

(4)LDA is computationally simpler (strong assumptions)than Bayes classifier

##### lda logistic and bayes decision

Practical models such as LDA, logistic regression that use assumptions and approximations to estimate decision boundaries based on data. They aim to approximate the theoretical bayes decision boundaries but are constrained by the limitations of the assumptions they make on the data.

### LDA example in the video

```{r}
# Load necessary libraries
library(MASS)  # For LDA
library(ROCR)  # For ROC curve and AUC
library(ggplot2)  # For plotting
# Load the ISLR package
library(ISLR)


# Load the Smarket dataset
data(Smarket)

# Set a seed for reproducibility
set.seed(123)

# Define the training set (using the data from 2001 to 2004)
train <- (Smarket$Year < 2005)

# Fit the LDA model using Lag1 and Lag2 as predictors
lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket, subset = train)

# Print LDA model summary
print(lda.fit)
lda.fit
plot(lda.fit)

# Make predictions on the 2005 data
lda.pred <- predict(lda.fit, Smarket[!train, ])

# Extract class predictions
lda.class <- lda.pred$class

# True labels for the test set
Direction.2005 <- Smarket$Direction[!train]

# Print the confusion matrix
conf_matrix <- table(lda.class, Direction.2005)
print(conf_matrix)

# Calculate and print the accuracy
lda_accuracy <- mean(lda.class == Direction.2005)
cat("Accuracy:", lda_accuracy, "\n")

# Print the posterior probabilities for the first 20 observations
cat("Posterior probabilities for the first 20 observations:\n")
print(lda.pred$posterior[1:20, 1])

# Print the class predictions for the first 20 observations
cat("Class predictions for the first 20 observations:\n")
print(lda.class[1:20])

# Count the number of observations with a posterior probability greater than 0.9
count_high_prob <- sum(lda.pred$posterior[, 1] > 0.9)
cat("Number of observations with a posterior probability greater than 0.9:", count_high_prob, "\n")

# Extract posterior probabilities for the positive class ('Up')
pred_prob <- lda.pred$posterior[, 2]  # Probabilities for the 'Up' class

# Create a binary vector for the actual class labels (1 for 'Up', 0 for 'Down')
true_labels_binary <- ifelse(Direction.2005 == "Up", 1, 0)

# Create a prediction object for ROC curve
pred_roc <- prediction(pred_prob, true_labels_binary)

# Create a performance object for ROC curve
perf_roc <- performance(pred_roc, measure = "tpr", x.measure = "fpr")

# Calculate AUC
auc <- performance(pred_roc, measure = "auc")
auc_value <- auc@y.values[[1]]
cat("AUC:", auc_value, "\n")

# Create a data frame for plotting the ROC curve
roc_data <- data.frame(
  FPR = unlist(perf_roc@x.values),
  TPR = unlist(perf_roc@y.values)
)

# Plot the ROC curve
roc_plot <- ggplot(roc_data, aes(x = FPR, y = TPR)) +
  geom_line(color = "blue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = paste("ROC Curve for LDA Model (AUC =", round(auc_value, 4), ")"),
       x = "False Positive Rate",
       y = "True Positive Rate") +
  theme_minimal()

# Print the ROC plot
print(roc_plot)

```

### the comparison between LDA and QDA in a classification

```{r}
# Load necessary libraries
library(MASS)  # For LDA and QDA models

# Set seed for reproducibility
set.seed(123)  # For reproducibility of the random sampling

# Create training and test sets
train <- sample(1:150, 120)  # Randomly select 120 samples for the training set out of 150

# Train LDA model
lda_iris <- lda(x = iris[, -5], grouping = iris[, 5], subset = train)  # Fit the LDA model using the training set
# Predict class labels for the test set
pred_class <- predict(lda_iris, iris[-train, -5])$class  # Predict the classes of the test set
# Generate confusion matrix for LDA
lda_conf_matrix <- table(Actual = iris[-train, 5], Predicted = pred_class)  # Create confusion matrix to compare actual vs. predicted classes
print("Confusion Matrix for LDA:")  # Print message
print(lda_conf_matrix)  # Display the confusion matrix
# Then there comes ROC curve but I could not draw it when it is not binary response
# Calculate and print LDA accuracy
lda_accuracy <- sum(diag(lda_conf_matrix)) / sum(lda_conf_matrix)  # Calculate accuracy as the proportion of correct classifications
lda_accuracy_rounded <- round(lda_accuracy, 4)  # Round accuracy to 4 decimal places
cat("Accuracy for LDA:", lda_accuracy_rounded, "\n")  # Print LDA accuracy

# Train QDA model
qda_iris <- qda(x = iris[, -5], grouping = iris[, 5], subset = train)  # Fit the QDA model using the training set
# Predict class labels for the test set
pred_class <- predict(qda_iris, iris[-train, -5])$class  # Predict the classes of the test set
# Generate confusion matrix for QDA
qda_conf_matrix <- table(Actual = iris[-train, 5], Predicted = pred_class)  # Create confusion matrix to compare actual vs. predicted classes
print("Confusion Matrix for QDA:")  # Print message
print(qda_conf_matrix)  # Display the confusion matrix

# Calculate and print QDA accuracy
qda_accuracy <- sum(diag(qda_conf_matrix)) / sum(qda_conf_matrix)  # Calculate accuracy as the proportion of correct classifications
qda_accuracy_rounded <- round(qda_accuracy, 4)  # Round accuracy to 4 decimal places
cat("Accuracy for QDA:", qda_accuracy_rounded, "\n")  # Print QDA accuracy

# Calculate log-likelihood for LDA and QDA
lda_logLik <- sum(log(predict(lda_iris, iris[-train, -5])$posterior[cbind(1:nrow(iris[-train, ]), as.numeric(iris[-train, 5]))]))  # Calculate log-likelihood for LDA
qda_logLik <- sum(log(predict(qda_iris, iris[-train, -5])$posterior[cbind(1:nrow(iris[-train, ]), as.numeric(iris[-train, 5]))]))  # Calculate log-likelihood for QDA

# Print log-likelihood values
cat("Log-Likelihood for LDA:", lda_logLik, "\n")  # Print LDA log-likelihood
cat("Log-Likelihood for QDA:", qda_logLik, "\n")  # Print QDA log-likelihood

# Perform Likelihood Ratio Test (LRT)
lrt_statistic <- -2 * (lda_logLik - qda_logLik)  # Compute the LRT statistic

# Degrees of freedom for the LRT test
df <- 2 * (ncol(iris) - 1) - 3  # Degrees of freedom: QDA has more parameters, so df = 2*number of features - parameters for LDA (LDA uses means and covariance)
p_value <- pchisq(lrt_statistic, df, lower.tail = FALSE)  # Compute p-value for the LRT

# Print LRT results
cat("Likelihood Ratio Test Statistic:", lrt_statistic, "\n")  # Print LRT statistic
cat("P-value for the Likelihood Ratio Test:", p_value, "\n")  # Print p-value for the LRT


```

### the comparison between the null model and LDA

```{r}
#### Iris Data
library(MASS)
library("ggthemes")
library("GGally")
library(hdrcde)
library(KernSmooth)
library("gridExtra")
library("vcd")
library(ggplot2)
data(iris)
head(iris)
ggpairs(iris,aes(col=Species, alpha=0.4))
```

```{r}
# LDA for iris data
train<-sample(1:150,120)
lda_iris<-lda(x=iris[train,-5],grouping=iris[train,5])
pred_class<-predict(lda_iris,iris[-train,-5])$class
### ???:$class
conf_matrix <- table(Actual = iris[-train, 5], Predicted = pred_class)
print(conf_matrix)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
cat("Accuracy:", accuracy, "\n")

# Function to calculate log-likelihood of LDA model
logLikLDA <- function(lda_model, data, actual_classes) {
  predicted <- predict(lda_model, data)
  posterior <- predicted$posterior
  log_lik <- sum(log(posterior[cbind(1:nrow(posterior), as.numeric(actual_classes))]))
  return(log_lik)
}

# Log-likelihood of the full model
logLik_full <- logLikLDA(lda_iris, iris[-train, -5], iris[-train, 5])

# Fit the reduced model (null model: only class proportions)
class_probs <- table(iris[train, 5]) / length(train)
logLik_null <- sum(log(class_probs[as.numeric(iris[-train, 5])]))

# Calculate the LRT statistic
LRT_statistic <- -2 * (logLik_null - logLik_full)
df <- length(levels(iris$Species)) - 1  # degrees of freedom

# Perform the test
p_value <- pchisq(LRT_statistic, df = df, lower.tail = FALSE)
cat("LRT Statistic:", LRT_statistic, "\n")
cat("Degrees of Freedom:", df, "\n")
cat("p-value:", p_value, "\n")

if (p_value < 0.05) {
  cat("The full model significantly improves the fit compared to the null model.\n")
} else {
  cat("The full model does not significantly improve the fit compared to the null model.\n")
}
### ?:\n
### why each time the accuracy(confution matrix is not the same??? maybe because of the random selection of training and testing data?)
```

### ROC=1.....

```{r}
# Load necessary libraries
library(MASS)  # For LDA
library(ROCR)  # For ROC curve and AUC
library(ggplot2)  # For plotting

# Load the iris dataset
data(iris)

# Filter the data to include only 'setosa' and 'versicolor'
iris_binary <- iris[iris$Species %in% c("setosa", "versicolor"), ]

# Convert 'Species' to a binary factor: 1 for 'setosa', 0 for 'versicolor'
iris_binary$Species_binary <- ifelse(iris_binary$Species == "setosa", 1, 0)

# Select only two variables for LDA
iris_binary <- iris_binary[, c("Sepal.Length", "Sepal.Width", "Species_binary")]

# Split the dataset into training (70%) and test (30%) sets
set.seed(123)  # For reproducibility
train_indices <- sample(1:nrow(iris_binary), 0.7 * nrow(iris_binary))
train_set <- iris_binary[train_indices, ]
test_set <- iris_binary[-train_indices, ]

# Train the LDA model on the binary classification problem
lda_model <- lda(Species_binary ~ Sepal.Length + Sepal.Width, data = train_set)

# Predict the class probabilities on the test set
predictions <- predict(lda_model, test_set[, c("Sepal.Length", "Sepal.Width")])
pred_prob <- predictions$posterior[, 2]  # Extract probabilities for the positive class ('setosa')

# Get true binary labels for the test set
true_labels_binary <- test_set$Species_binary

# Create a prediction object for ROC curve
pred_roc <- prediction(pred_prob, true_labels_binary)

# Create a performance object for ROC curve
perf_roc <- performance(pred_roc, measure = "tpr", x.measure = "fpr")

# Calculate AUC
auc <- performance(pred_roc, measure = "auc")
auc_value <- auc@y.values[[1]]
cat("AUC:", auc_value, "\n")

# Create a data frame for plotting the ROC curve
roc_data <- data.frame(
  FPR = unlist(perf_roc@x.values),
  TPR = unlist(perf_roc@y.values)
)

# Plot the ROC curve
roc_plot <- ggplot(roc_data, aes(x = FPR, y = TPR)) +
  geom_line(color = "blue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = paste("ROC Curve for LDA Model (AUC =", round(auc_value, 4), ")"),
       x = "False Positive Rate",
       y = "True Positive Rate") +
  theme_minimal()

# Print the ROC plot
print(roc_plot)

```

### the comparison between Naive bayes and LDA

```{r}
# Load necessary libraries
library(e1071)  # For naiveBayes
library(MASS)   # For lda
library(datasets)

# Load the Iris dataset
data(iris)

# Set a seed for reproducibility
set.seed(123)

# Split the dataset into training and test sets (70% train, 30% test)
sample_indices <- sample(1:nrow(iris), size = 0.7 * nrow(iris))
train_set <- iris[sample_indices, ]
test_set <- iris[-sample_indices, ]

# Train the Naive Bayes model
nb_model <- naiveBayes(Species ~ ., data = train_set)

# Make predictions on the test set
nb_predictions <- predict(nb_model, test_set)
nb_conf_matrix <- table(Predicted = nb_predictions, Actual = test_set$Species)

# Print the confusion matrix for Naive Bayes
cat("Confusion Matrix for Naive Bayes:\n")
print(nb_conf_matrix)

# Calculate and print the accuracy for Naive Bayes
nb_accuracy <- sum(diag(nb_conf_matrix)) / sum(nb_conf_matrix)
cat("Accuracy for Naive Bayes:", round(nb_accuracy, 4), "\n")

# Train the LDA model
lda_model <- lda(Species ~ ., data = train_set)

# Make predictions on the test set
lda_predictions <- predict(lda_model, test_set)$class
lda_conf_matrix <- table(Predicted = lda_predictions, Actual = test_set$Species)

# Print the confusion matrix for LDA
cat("\nConfusion Matrix for LDA:\n")
print(lda_conf_matrix)

# Calculate and print the accuracy for LDA
lda_accuracy <- sum(diag(lda_conf_matrix)) / sum(lda_conf_matrix)
cat("Accuracy for LDA:", round(lda_accuracy, 4), "\n")

# Compute Log-Likelihood for Naive Bayes
nb_logLik <- sum(log(predict(nb_model, test_set, type = "raw")[cbind(1:nrow(test_set), as.numeric(test_set$Species))]))
cat("\nLog-Likelihood for Naive Bayes:", round(nb_logLik, 4), "\n")

# Compute Log-Likelihood for LDA
lda_logLik <- sum(log(predict(lda_model, test_set)$posterior[cbind(1:nrow(test_set), as.numeric(test_set$Species))]))
cat("Log-Likelihood for LDA:", round(lda_logLik, 4), "\n")

# Perform the Likelihood Ratio Test
lrt_statistic <- -2 * (lda_logLik - nb_logLik)

# Compute degrees of freedom for the LRT
df <- 0  # Degrees of freedom for LRT between LDA and Naive Bayes is typically considered 0 in this case (complex models often require detailed adjustment)

# Compute p-value for the LRT
p_value <- pchisq(lrt_statistic, df, lower.tail = FALSE)
cat("Likelihood Ratio Test Statistic:", round(lrt_statistic, 4), "\n")
cat("P-value for the Likelihood Ratio Test:", round(p_value, 4), "\n")

```

### the comparison between LDA and logistic regression

```{r}
# Load necessary libraries
library(MASS)
library(glmnet)
library(ggplot2)
library(gridExtra)

# Subset the data to exclude 'virginica' for binary classification
iris_binary <- subset(iris, Species != "virginica")
iris_binary$Species <- factor(iris_binary$Species)

# Prepare data for glmnet
x <- as.matrix(iris_binary[, -5])
y <- iris_binary$Species

# Fit regularized logistic regression model
logistic_model <- cv.glmnet(x, y, family = "binomial", alpha = 0.5) # alpha=0.5 for elastic net

# Fit LDA model
lda_model <- lda(Species ~ ., data = iris_binary)

# Predict probabilities from logistic regression
logistic_pred <- predict(logistic_model, newx = x, type = "response", s = "lambda.min")
logistic_class <- ifelse(logistic_pred > 0.5, "versicolor", "setosa")

# Predict probabilities from LDA
lda_pred <- predict(lda_model)$posterior
lda_class <- ifelse(lda_pred[, "versicolor"] > 0.5, "versicolor", "setosa")

# Confusion matrices for both models
logistic_conf_matrix <- table(Actual = iris_binary$Species, Predicted = logistic_class)
lda_conf_matrix <- table(Actual = iris_binary$Species, Predicted = lda_class)

# Print confusion matrices
cat("Confusion Matrix for Logistic Regression:\n")
print(logistic_conf_matrix)

cat("\nConfusion Matrix for LDA:\n")
print(lda_conf_matrix)

# Calculate and print accuracies
logistic_accuracy <- sum(diag(logistic_conf_matrix)) / sum(logistic_conf_matrix)
lda_accuracy <- sum(diag(lda_conf_matrix)) / sum(lda_conf_matrix)

cat("Accuracy for Logistic Regression:", logistic_accuracy, "\n")
cat("Accuracy for LDA:", lda_accuracy, "\n")

# Log-likelihood for logistic regression
logistic_logLik <- sum(dbinom(as.numeric(iris_binary$Species == "versicolor"), size = 1, prob = logistic_pred, log = TRUE))
cat("Log-Likelihood for Logistic Regression:", logistic_logLik, "\n")

# Log-likelihood for LDA
lda_logLik <- sum(log(lda_pred[cbind(1:nrow(iris_binary), as.numeric(iris_binary$Species))]))
cat("Log-Likelihood for LDA:", lda_logLik, "\n")

# LRT statistic
lrt_statistic <- -2 * (logistic_logLik - lda_logLik)
cat("Likelihood Ratio Test Statistic:", lrt_statistic, "\n")

# Degrees of freedom for the test
df <- length(coef(logistic_model)) - 1 # Adjusted degrees of freedom for regularized model

# P-value for the test
p_value <- pchisq(lrt_statistic, df, lower.tail = FALSE)
cat("P-value for the Likelihood Ratio Test:", p_value, "\n")

# Create a grid of values for plotting decision boundaries
plot_grid <- expand.grid(
  Sepal.Length = seq(min(iris_binary$Sepal.Length), max(iris_binary$Sepal.Length), length.out = 100),
  Sepal.Width = seq(min(iris_binary$Sepal.Width), max(iris_binary$Sepal.Width), length.out = 100),
  Petal.Length = mean(iris_binary$Petal.Length),  # Use mean value for Petal.Length
  Petal.Width = mean(iris_binary$Petal.Width)    # Use mean value for Petal.Width
)
plot_grid_matrix <- as.matrix(plot_grid)

# Predict class probabilities for the grid points using logistic regression
logistic_pred_grid <- predict(logistic_model, newx = plot_grid_matrix, type = "response", s = "lambda.min")
logistic_class_grid <- ifelse(logistic_pred_grid > 0.5, "versicolor", "setosa")

# Predict class probabilities for the grid points using LDA
lda_pred_grid <- predict(lda_model, newdata = plot_grid)$posterior
lda_class_grid <- ifelse(lda_pred_grid[, "versicolor"] > 0.5, "versicolor", "setosa")

# Convert the grid predictions to data frames for plotting
logistic_plot_grid <- cbind(plot_grid[, 1:2], Predicted = factor(logistic_class_grid, levels = c("setosa", "versicolor")))
lda_plot_grid <- cbind(plot_grid[, 1:2], Predicted = factor(lda_class_grid, levels = c("setosa", "versicolor")))

# Plot for logistic regression
logistic_plot <- ggplot() +
  geom_point(data = iris_binary, aes(x = Sepal.Length, y = Sepal.Width, color = Species), size = 2) +
  geom_tile(data = logistic_plot_grid, aes(x = Sepal.Length, y = Sepal.Width, fill = Predicted), alpha = 0.3) +
  scale_color_manual(values = c("setosa" = "red", "versicolor" = "blue")) +
  scale_fill_manual(values = c("setosa" = "red", "versicolor" = "blue")) +
  labs(title = "Logistic Regression Classification Boundaries") +
  theme_minimal()

# Plot for LDA
lda_plot <- ggplot() +
  geom_point(data = iris_binary, aes(x = Sepal.Length, y = Sepal.Width, color = Species), size = 2) +
  geom_tile(data = lda_plot_grid, aes(x = Sepal.Length, y = Sepal.Width, fill = Predicted), alpha = 0.3) +
  scale_color_manual(values = c("setosa" = "red", "versicolor" = "blue")) +
  scale_fill_manual(values = c("setosa" = "red", "versicolor" = "blue")) +
  labs(title = "LDA Classification Boundaries") +
  theme_minimal()

# Arrange the two plots side by side
grid.arrange(logistic_plot, lda_plot, ncol = 2)

```

```{r}
# Check for perfect separation
plot(iris_binary$Sepal.Length, iris_binary$Sepal.Width, col = iris_binary$Species)

library(glmnet)

# Prepare data for glmnet
x <- as.matrix(iris_binary[, -5])
y <- iris_binary$Species

# Fit regularized logistic regression model
logistic_model <- cv.glmnet(x, y, family = "binomial", alpha = 0.5) # alpha=0.5 for elastic net
logistic_model
# Fit LDA model
lda_model <- lda(Species ~ ., data = iris_binary)
lda_model
# Predict probabilities from logistic regression
logistic_pred <- predict(logistic_model, newx = x, type = "response", s = "lambda.min")
logistic_class <- ifelse(logistic_pred > 0.5, "versicolor", "setosa")

# Predict probabilities from LDA
lda_pred <- predict(lda_model)$posterior
lda_class <- ifelse(lda_pred[, "versicolor"] > 0.5, "versicolor", "setosa")

# Confusion matrices for both models
logistic_conf_matrix <- table(Actual = iris_binary$Species, Predicted = logistic_class)
lda_conf_matrix <- table(Actual = iris_binary$Species, Predicted = lda_class)

# Print confusion matrices
cat("Confusion Matrix for Logistic Regression:\n")
print(logistic_conf_matrix)

cat("\nConfusion Matrix for LDA:\n")
print(lda_conf_matrix)

# Calculate and print accuracies
logistic_accuracy <- sum(diag(logistic_conf_matrix)) / sum(logistic_conf_matrix)
lda_accuracy <- sum(diag(lda_conf_matrix)) / sum(lda_conf_matrix)

cat("Accuracy for Logistic Regression:", logistic_accuracy, "\n")
cat("Accuracy for LDA:", lda_accuracy, "\n")
# Log-likelihood for logistic regression
logistic_logLik <- sum(dbinom(as.numeric(iris_binary$Species == "versicolor"), size = 1, prob = logistic_pred, log = TRUE))
cat("Log-Likelihood for Logistic Regression:", logistic_logLik, "\n")

# Log-likelihood for LDA
lda_logLik <- sum(log(lda_pred[cbind(1:nrow(iris_binary), as.numeric(iris_binary$Species))]))
cat("Log-Likelihood for LDA:", lda_logLik, "\n")

# LRT statistic
lrt_statistic <- -2 * (logistic_logLik - lda_logLik)
cat("Likelihood Ratio Test Statistic:", lrt_statistic, "\n")

# Degrees of freedom for the test
df <- length(coef(logistic_model)) - 1 # Adjusted degrees of freedom for regularized model

# P-value for the test
p_value <- pchisq(lrt_statistic, df, lower.tail = FALSE)
cat("P-value for the Likelihood Ratio Test:", p_value, "\n")


```

# kNN

```{r}
library(class)
# Bank Data
data(bank,package="gclus")
head(bank)
bank$Status<-factor(bank$Status)

ggpairs(bank, aes(alpha=0.4))
ggpairs(bank,aes(col=Status, alpha=0.4))

test_bnk <- seq(from=1,to=nrow(bank),by=4)
train_bnk <- c(1:nrow(bank))[-test_bnk]

train.set <- bank[train_bnk,]
test.set <- bank[test_bnk,]
labels <- factor(x=bank[train_bnk,1])

# kNN for k = 1
output <- knn(train=train.set[,-1], test=test.set[,-1], cl=labels, k=1, prob=TRUE) 
table(output,test.set[,1])

# kNN for k = 1,3,5,7,9
for(k in c(1,3,5,7,9)){
	cat(paste("There are",k,"nearest neighbours"),"\n")
	output <- knn(train=train.set[,-1], test=test.set[,-1], cl=labels, k=k, prob=TRUE)
	print(output)
	print(table(output,test.set[,1]))
}
```

```{r}

### Wine data (p=13) from gclus
data(wine,package="gclus")
head(wine)
wine$Class<-factor(wine$Class)
ggpairs(wine,aes(col=Class, alpha=0.4))

s.wine <- scale(wine[,-1])
s.wine <- cbind(wine[,1],s.wine)

train.vals <- sample(c(1:nrow(s.wine)),120)

train.set <- s.wine[train.vals,]
test.set <- s.wine[-train.vals,]
labels <- factor(x=s.wine[train.vals,1])

prop <- 0.8
validation.set <- sample(c(1:nrow(train.set)),size=prop*length(train.vals))
labels.validation.set <- factor(x=train.set[validation.set,1])

for(k in c(1,3,5,7,9,11,13,15)){
	cat(paste("There are",k,"nearest neighbours in the validation set"),"\n")
	output <- knn(train=train.set[validation.set,-1], test=train.set[-validation.set,-1], cl=labels.validation.set, k=k, prob=TRUE)
	print(output)
	class.table <- table(output,train.set[-validation.set,1])
	print(class.table)
	print(sum(diag(class.table))/sum(class.table))
}
	
for(k in c(7,9,11)){
	cat(paste("There are",k,"nearest neighbours"),"\n")
	output <- knn(train=train.set[,-1], test=test.set[,-1], cl=labels, k=k, prob=TRUE)
	print(output)
	class.table <- table(output,test.set[,1])
	print(class.table)
	print(sum(diag(class.table))/sum(class.table))
}
	
# And this is what happens if the data are not scaled...
train.vals <- sample(c(1:nrow(wine)),120)

train.set <- wine[train.vals,]
test.set <- wine[-train.vals,]
labels <- factor(x=wine[train.vals,1])

prop <- 0.8
validation.set <- sample(c(1:nrow(train.set)),size=prop*length(train.vals))
labels.validation.set <- factor(x=train.set[validation.set,1])

for(k in c(1,3,5,7,9,11,13,15)){
	cat(paste("There are",k,"nearest neighbours in the validation set"),"\n")
	output <- knn(train=train.set[validation.set,-1], test=train.set[-validation.set,-1], cl=labels.validation.set, k=k, prob=TRUE)
	print(output)
	class.table <- table(output,train.set[-validation.set,1])
	print(class.table)
	print(sum(diag(class.table))/sum(class.table))
}
	
for(k in c(9)){
	cat(paste("There are",k,"nearest neighbours"),"\n")
	output <- knn(train=train.set[,-1], test=test.set[,-1], cl=labels, k=k, prob=TRUE)
	print(output)
	class.table <- table(output,test.set[,1])
	print(class.table)
	print(sum(diag(class.table))/sum(class.table))
}

```
